Structured Versus Decorated Cospans

### 2021-01-08 cospans.tex
"much simpler proof" |-> "alternative proof"
various small edits and comments, finished my review

There is a much simpler proof of the above theorem when the hypothesis of \cref{thm:equiv} hold.  In this case we show, in the proof of \cref{thm:equiv}, that the decorated cospan double category $F \lCsp$ is isomorphic to a certain double category ${}_L \lCsp(\inta  F)$ constructed using \emph{structured} cospans.   We also show that under this isomorphism, all the symmetric monoidal structure of $F \lCsp$---the tensor product, associator, etc.---matches that for  ${}_L \lCsp(\inta  F)$, which we have proved to obey the axioms of a symmetric monoidal double category (see \cref{thm:SC}).  It follows that $F \lCsp$ is a symmetric monoidal double category.

@@

As a left adjoint, this induced $L$ preserves all colimits that exist between the categories $\A$ and $\inta F$, which have finite colimits, so we can construct the double category of structured cospans $_L\lCsp(\inta F)$ of \cref{thm:SC}. Its objects and vertical morphisms are those of the category $\A$ (just as for $F\lCsp$), whereas its horizontal 1-cells are now cospans of the form $\cspn{L(a)}{(m,x)}{L(b)}{}{}$ in the Grothendieck category $\inta F$ (see \cref{def:GrothCat}). Explicitly, they consist of two pairs of morphisms
\begin{equation}\label{eq:scsphor1cell}
 (a,\bot_a)\xrightarrow{\scalebox{0.7}{\(\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}\)}}(m,x)\xleftarrow{\scalebox{0.7}{\(\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}\)}}(b,\bot_b)
\end{equation}
with $x\in F(m)$; recall that the reindexing functors $F(i),F(o)$ preserve all finite colimits across the fibers.
Finally, the 2-morphisms of this double category in this context unravel as follows:
\begin{equation}\label{eq:2morphismLCspintaF}
 \begin{tikzcd}[row sep=1in,column sep=1.7in,ampersand replacement=\&]
 (a,\bot_a)\ar[r,"{\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}}"]\ar[d,"{\begin{cases}f\maps a\to a' &\textrm{in }\A \\\chi_a\maps F(f)(\bot_a)\cong \bot_{a'} &\textrm{in }F(a')\end{cases}}"description] \& (m,x) \ar[d,"{\begin{cases}k\maps m\to m' &\textrm{in }\A \\ \tau\maps F(k)(x)\to x' &\textrm{in }F(m')\end{cases}}"description] \& (b,\bot_b)\ar[l,"{\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}}"']\ar[d,"{\begin{cases}g\maps b\to b' &\textrm{in }\A \\\chi_b\maps F(g)(\bot_b)\cong \bot_{b'} &\textrm{in }F(b')\end{cases}}"description] \\
 (a',\bot_{a'})\ar[r,"{\begin{cases}i'\maps a'\to m' &\textrm{in }\A \\!\maps F(i')(\bot_{a'})\to x' &\textrm{in }F(m')\end{cases}}"'] \& (m',x') \& (b',\bot_{b'})\ar[l,"{\begin{cases}o'\maps b'\to m' &\textrm{in }\A \\!\maps F(o')(\bot_{b'})\to x' &\textrm{in }F(m')\end{cases}}"]
 \end{tikzcd}
\end{equation}
where the outside vertical legs come from the definition of $L$ on arrows \cref{eq:Lonarrows}. 
Explicitly, the commutativity of the above squares translates to the conditions $k\circ i=i'\circ f$ and $k\circ o=o'\circ g$ in $\A$, and also to
\begin{gather}\label{eq:Grothcommutativity}
 F(k\circ i)(\bot_a)\cong Fk(Fi(\bot_a))\xrightarrow{Fk(!)}Fk(x)\xrightarrow{\tau}x'= \\
 F(i'\circ f)(\bot_a)\cong Fi'(Ff(\bot_a))\xrightarrow{Fi'(\chi_a)}Fi'(\bot_{a'})\xrightarrow{!}x'\nonumber
\end{gather}
in the fiber $F(m')$, by composition in the Grothendieck category \cref{eq:compGrothcat}. Since both morphisms emanate from the mapped initial object in $F(m')$ under the finite-colimit-preserving reindexing functors, they are the unique such into $x'$ therefore this gives no extra conditions; similarly for the equality including $o,o'$.

Building towards an isomorphism of the above described double categories, we now define a double functor sending structured cospans to decorated cospans:
\begin{displaymath}
\lE=(\lE_0,\lE_1)\maps _L\lCsp(\inta F) \longrightarrow F\lCsp.
\end{displaymath}
We set $\lE_0 \maps \A \to \A$ to be the identity, and decree that $\lE_1$ sends any
$L$-structured cospan as in \cref{eq:scsphor1cell} to the decorated cospan
$a \xrightarrow{i} m \xleftarrow{o} b$ with decoration $x\in F(m)$, whereas $\lE_1$ sends 
any map of $L$-structured cospans as in \cref{eq:2morphismLCspintaF} to the
map of $F$-decorated cospans given by
\[
\begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"k"] & b\ar[l,"o"']\ar[d,"g"] \\
a'\ar[r,"i'"']& m' & b'\ar[l,"o'"]
\end{tikzcd}
\]
along with the morphism of decorations $\tau\maps F(k)(x) \to x'$ in $F(m')$ coming from the middle arrow above. We then make $\lE$ into double functor as in \cref{def:doublefun}. First, both $\lE_0$ and $\lE_1$ are functors that appropriately commute with sources and targets.  Second, we equip $\lE$ with globular 2-isomorphisms $\lE(N) \odot \lE(M) \simrightarrow\lE(N \odot M)$, $U_{\lE m} \simrightarrow\lE(U_m)$
for any pair of composable horizontal 1-cells $M=\cspn{(a,\bot_a)}{(m,x)}{(b,\bot_b)}{}{}$ and $N=\cspn{(b,\bot_b)}{(n,y)}{(c,\bot_c)}{}{}$ and any object $m$ of $_L\lCsp(\inta F)$. Explicitly, $\mathbb{E}(N)\odot \mathbb{E}(M)$ is given as in \cref{eq:dcospanscomposition} via the pushout cospan and decoration
\begin{equation}\label{eq:1}
 \begin{tikzcd}
  & m+_bn & \\
  a\ar[ur] && b\ar[ul]
 \end{tikzcd}\qquad
  \one\xrightarrow{x\times y}F(m)\times F(n)\xrightarrow{\phi_{m,n}}F(m+n)\xrightarrow{F(\psi)}F(m+_bn).
\end{equation}
On the other hand, $N \odot M$ computed as in \cref{eq:globalpushout} is 
\begin{displaymath}
 \begin{tikzcd}[column sep={.7in,between origins}]
 && (m+_b n,F(v_m)x+_{(\bot_{m+_bn})}F(v_n)y) && \\
 & (m,x) \ar[ur] && (n,y)\ar[ul] \\
 (a,\bot_a)\ar[ur] && (b,\bot_b)\ar[ur]\ar[ul] && (c,\bot_c)\ar[ul]
 \end{tikzcd}
\end{displaymath}
where $v_m,v_n$ are the canonical maps into the pushout.  Under $\lE$ this is mapped to the cospan $\cspn{a}{m+_bn}{b}{}{}$ with decoration $F(v_m)x+F(v_n)y$ in the fiber $F(m+_bn)$.  This decoration is canonically isomorphic to that of \cref{eq:1} due to the following diagram 
\begin{equation}\label{eq:isomorphismdiag}
 \begin{tikzcd}[sep=.4in]
F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[d,"F(v_m)\times F(v_n)"']\ar[dr,phantom,"\stackrel{\phi_{v_m,v_n}}{\cong}"] & F(m+n)\ar[r,"F(\psi)"]\ar[dr,phantom,"\cong"']\ar[d,"F(v_m+v_n)"] & F(m+_bn)\ar[d,"F(1)"]\\
F(m+_bn)\times F(m+_bn)\ar[r,"\phi_{m+_bn,m+_bn}"'] & F\left((m+_bn)+(m+_bn)\right)\ar[r,"F(\nabla)"'] & F(m+_bn)
 \end{tikzcd}
\end{equation}
where the composite of the bottom two arrows is the coproduct $+ \maps F(m+_bn) \times F(m+_bn) \to F(m+_bn)$ as in \cref{eq:explicitstructure1}. The left-hand isomorphism comes from pseudonaturality of $\phi$ as in \cref{eq:pseudonaturality}, whereas the right-hand isomorphism follows from universal properties and pseudofunctoriality of $F$.  Thus, we have a globular 2-isomorphism  $\lE(N) \odot \lE(M) \simrightarrow\lE(N \odot M)$. Notice that since the target decoration is part of the above pushout in the Grothendieck category $\inta F$, this isomorphism we construced is the canonical such that it commutes with the respective cocones. This is useful when checking associativity for example: there is only one such canonical isomorphism $\lE(P)\odot\lE(N)\odot\lE(M)\to\lE(P\odot N\odot M)$ for any three composable structured cospans.

Regarding identities, the identity horizontal 1-cell $U_m$ in $_L\lCsp(\inta F)$ is $\cspn{(m,\bot_m)}{(m,\bot_m)}{(m,\bot_m)}{}{}$ with $1_m$ as the $A$-component and isomorphisms between initial objects in the fibers. Thus $\lE(U_m)$ is the identity cospan in $A$ together with the `initial decoration' $\bot_m\in F(m)$, and so is $U_{\mathbb{E}(m)}$ as in \cref{eq:UFCsp}.

@@

Since both double categories are symmetric monoidal under the running assumptions (\cref{DC,thm:SC}), we now show that this double isomorphism is a symmetric monoidal one, namely that $\lE=(\lE_0,\lE_1)$ is a symmetric strong monoidal double functor as per \cref{defn:monoidal_double_functor}. Indeed, $\lE_0= \id_\A$ is a symmetric monoidal functor trivially. For $\lE_1$, given two structured cospans $M=(a,\bot_a)\to(m,x)\leftarrow(b,\bot_b)$ and $M'=(a',\bot_{a'})\to(m',x')\leftarrow(b',\bot_{b'})$ in $_L\lCsp(\inta F)$, we compute the appropriate decorated cospans
\begin{align*}
\lE(M)\ot\lE(M')&\;\stackrel{\cref{eq:tensordecoration}}{=}\;a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\phi_{m,m'}(x,x')\in F(m+m')
%\one\xrightarrow{x\times x'}F(m)\times F(m')\xrightarrow{\phi_{m,m'}}F(m+m')
\\
\lE(M\ot M')&\;\stackrel{\cref{eq:tensorstuctured}}{=}\;\lE\Big((a+a',\bot_{a+a'})\to\underbrace{(m+m',F(\iota_m)x+F(\iota_{m'})x')}_{\cref{eq:fibrewisecoprod}}\leftarrow(b+b',\bot_{b+b'})\Big) \\
&=a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }F(\iota_m)x+F(\iota_{m'})x'\in F(m+m')
\end{align*}
The cospan part is equal, whereas we have an isomorphism between the decorations, $F(\iota_m)x+F(\iota_{m'})x'\cong \phi_{m,m'}(x,x')$, given in an analogous way to \cref{eq:isomorphismdiag} as follows:
\begin{displaymath}
\begin{tikzcd}[sep=.4in]
F(m)\times F(m')\ar[r,"\phi_{m,m'}"]\ar[d,"F(\iota_m)\times F(\iota_{m'})"']
\ar[dr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{\iota_m,\iota_n}}{\cong}}"] & F(m+m')\ar[dr,"F(1)"]\ar[d,bend left=60, phantom,"\cong"']\ar[d,"F(\iota_m+\iota_{m'})"']  & \\
F(m+m')\times F(m+m')\ar[r,"\phi_{m+m',m+m'}"'] & F\left((m+m')+(m+m')\right)\ar[r,"F(\nabla)"'] & F(m+m')
 \end{tikzcd}
\end{displaymath}
where the composite of the bottom two arrows is the coproduct operation in $F(m+m')$.
We thus have constructed an isomorphism $\lE(M)\otimes\lE(M') \simrightarrow \lE(M\otimes M')$ for each $M,M'$. 
Again, notice that since the target decoration is part of the sum $\left(m+m',F(\iota_m)x+F(\iota_{m'})x'\right)$ in the Grothendieck category $\inta F$ by \cref{eq:fibrewisecoprod}, this isomorphism we constructured is the canonical such that it commutes with the respective inclusions into the sum, and there is only one such canonical isomorphism $\lE(M)\otimes\lE(M')\otimes\lE(M'')\simrightarrow\lE(M\ot M'\ot M'')$ for any three structured cospans, proving associativity.

@@

\begin{proof}
This follows by applying Shulman's result \cite[Theorem 1.2]{Shulman2010} to the fibrant symmetric monoidal double category $F\lCsp$.
\end{proof}

@@

\begin{thm} \label{thm:bicat_equiv}
Suppose $\A$ has finite colimits and $F \maps(\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor that factors through $\Rex$ as an ordinary pseudofunctor.    Define the symmetric monoidal bicategory $_L\bCsp(\inta F)$ as in \cref{thm:equiv}.   Then there is an isomorphism of symmetric monoidal bicategories
\[      F\bCsp \simeq {}_L \bCsp(\inta F)   \]
and of symmetric monoidal categories
\[      F\Csp \simeq {}_L \Csp(\inta F)  . \]
\end{thm}

\begin{proof} 
Hansen and Shulman \cite{HS} showed that the passage from symmetric monoidal double categories to symmetric monoidal bicategories is functorial in a suitable sense.  This implies that an isomorphism of symmetric monoidal double categories $\lD \simeq \lD'$ gives an isomorphism of symmetric monoidal bicategories $\bD \simeq \bD'$.    Since the process of decategorifying a bicategory merely discards 2-morphisms and takes isomorphism classes of 1-morphisms, the isomorphism of symmetric monoidal bicategories $\bD \simeq \bD'$ in turn induces an isomorphism of symmetric monoidal categories $\D \simeq \D'$.   Thus, the theorem follows from \cref{thm:equiv}. \end{proof}

@@

We thus have two symmetric monoidal double categories: ${}_L \lCsp(\Graph)$ obtained from structured cospans and $F\lCsp$ obtained from decorated cospans. Each of these double categories has $\Fin\Set$ as its category of objects, open graphs as horizontal 1-cells, and maps of open graphs as 2-morphisms.   This suggests that ${}_L \lCsp(\Graph)$  and $F\lCsp$ are isomorphic as symmetric monoidal double categories---and indeed this is true.   This follows from \cref{thm:equiv} together with the fact that $U \maps \inta F \to \Fin\Set$ and $R \maps \Graph \to \Fin\Set$ are isomorphic as symmetric monoidal opfibrations.

@@

Given a Petri net with rates $P$ and defining $v$ as above, this system of first-order ordinary differential equations for a function $c \maps \R \to \R^S$:
\[    \frac{d}{dt} c(t) = v(c(t))   \]
is called the \define{rate equation}.    When $P$ is part of an open Petri net with rates
\[
\begin{tikzpicture}[scale=1.5]
\node (D) at (-3,0) {$L(X)$};
\node (E) at (-2,0) {$P$};
\node (F) at (-1,0) {$L(Y)$,};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$i$} (E)
(F) edge node [above] {$o$} (E);
\end{tikzpicture}
\]
we get an open dynamical system called the \define{open rate equation}:
\[     \frac{d}{dt} c(t) = v(c(t))  + i_*(I(t)) - o_*(O(t)) \]
where $I \maps \R \to \R^X$ and $O \maps \R \to \R^X$ are arbitrary smooth functions describing \define{inflows} and \define{outflows}, respectively.   Applying this prescription to the open Petri net with rates shown at the start of this section one gets the differential equations \cref{eq:openPetrir}.  Other examples are worked out in \cite{BP}.

We now show that decorated cospan double category $D \lCsp$ of open dynamical systems is not isomorphic to a structured cospan double category via \cref{thm:equiv}.  Recall that in that theorem we start with the data required to build a decorated cospan category, namely a symmetric lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat,\times)$, and show that under certain conditions the resulting opfibration $U \maps \X = \inta F \to \A$ has a left adjoint $L \maps \A \to \X$.   We then obtain an isomorphism between decorated and structured cospan double categories, $F \lCsp \simeq {}_L \Csp(\X)$.   We now show that in the case at hand, where $F = D$,  the opfibration $U$ does \emph{not} have a left adjoint.  Thus, the conditions of \cref{thm:equiv} do not hold in this case.

@@

where the unnamed isomorphisms are appropriate composites of the pseudoassociativity of $F$ \cref{eq:omega}. The result now follows from lengthy pasted diagram calculations using the axioms of a sylleptic lax monoidal pseudofunctor.

>>>>

There is an alternative proof of the above theorem when the hypothesis of \cref{thm:equiv} hold.  In this case we show, in the proof of \cref{thm:equiv}, that the decorated cospan double category $F \lCsp$ is isomorphic to a certain double category ${}_L \lCsp(\inta  F)$ constructed using \emph{structured} cospans.   We also show that under this isomorphism, all the symmetric monoidal structure of $F \lCsp$---the tensor product, associator, etc.---matches that for  ${}_L \lCsp(\inta  F)$, which we have proved to obey the axioms of a symmetric monoidal double category (see \cref{thm:SC}).  It follows that $F \lCsp$ is a symmetric monoidal double category.

@@

As a left adjoint, this induced $L$ preserves all colimits that exist between the categories $\A$ and $\inta F$, which have finite colimits, so we can construct the double category of structured cospans $_L\lCsp(\inta F)$ of \cref{thm:SC}. Its objects and vertical morphisms are those of the category $\A$ (just as for $F\lCsp$), whereas its horizontal 1-cells are now cospans of the form $\cspn{L(a)}{(m,x)}{L(b)}{}{}$ in the Grothendieck category $\inta F$ (see \cref{def:GrothCat}). Explicitly, they consist of two pairs of morphisms
\begin{equation}\label{eq:scsphor1cell}
 (a,\bot_a)\xrightarrow{\scalebox{0.7}{\(\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}\)}}(m,x)\xleftarrow{\scalebox{0.7}{\(\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}\)}}(b,\bot_b)
\end{equation}
with $x\in F(m)$; recall that the reindexing functors $F(i),F(o)$ preserve all finite colimits across the fibers.
Finally, the 2-morphisms of this double category in this context unravel as follows:
\begin{equation}\label{eq:2morphismLCspintaF}
\scalebox{.9}
 {\begin{tikzcd}[row sep=1in,column sep=1.7in,ampersand replacement=\&]
 (a,\bot_a)\ar[r,"{\begin{cases}i\maps a\to m &\textrm{in }\A \\!\maps F(i)(\bot_a)\to x &\textrm{in }F(m)\end{cases}}"]\ar[d,"{\begin{cases}f\maps a\to a' &\textrm{in }\A \\\chi_a\maps F(f)(\bot_a)\cong \bot_{a'} &\textrm{in }F(a')\end{cases}}"description] \& (m,x) \ar[d,"{\begin{cases}k\maps m\to m' &\textrm{in }\A \\ \tau\maps F(k)(x)\to x' &\textrm{in }F(m')\end{cases}}"description] \& (b,\bot_b)\ar[l,"{\begin{cases}o\maps b\to m &\textrm{in }\A \\!\maps F(o)(\bot_b)\to x &\textrm{in }F(m)\end{cases}}"']\ar[d,"{\begin{cases}g\maps b\to b' &\textrm{in }\A \\\chi_b\maps F(g)(\bot_b)\cong \bot_{b'} &\textrm{in }F(b')\end{cases}}"description] \\
 (a',\bot_{a'})\ar[r,"{\begin{cases}i'\maps a'\to m' &\textrm{in }\A \\!\maps F(i')(\bot_{a'})\to x' &\textrm{in }F(m')\end{cases}}"'] \& (m',x') \& (b',\bot_{b'})\ar[l,"{\begin{cases}o'\maps b'\to m' &\textrm{in }\A \\!\maps F(o')(\bot_{b'})\to x' &\textrm{in }F(m')\end{cases}}"]
 \end{tikzcd}}
\end{equation}
where the outside vertical legs come from the definition of $L$ on arrows \cref{eq:Lonarrows}. 
Explicitly, the commutativity of the above squares translates to the conditions $k\circ i=i'\circ f$ and $k\circ o=o'\circ g$ in $\A$, and also to
\begin{gather}\label{eq:Grothcommutativity}
 F(k\circ i)(\bot_a)\cong Fk(Fi(\bot_a))\xrightarrow{Fk(!)}Fk(x)\xrightarrow{\tau}x'= \\
 F(i'\circ f)(\bot_a)\cong Fi'(Ff(\bot_a))\xrightarrow{Fi'(\chi_a)}Fi'(\bot_{a'})\xrightarrow{!}x'\nonumber
\end{gather}
in the fiber $F(m')$, by composition in the Grothendieck category \cref{eq:compGrothcat}. Since both morphisms emanate from the mapped initial object in $F(m')$ under the finite-colimit-preserving reindexing functors, they are the unique such into $x'$ therefore this gives no extra conditions; similarly for the equality including $o,o'$.

Building towards an isomorphism of the above described double categories, we now define a double functor sending structured cospans to decorated cospans:
\begin{displaymath}
\lE=(\lE_0,\lE_1)\maps _L\lCsp(\inta F) \longrightarrow F\lCsp.
\end{displaymath}
We set $\lE_0 \maps \A \to \A$ to be the identity, and decree that $\lE_1$ sends any
$L$-structured cospan as in \cref{eq:scsphor1cell} to the decorated cospan
$a \xrightarrow{i} m \xleftarrow{o} b$ with decoration $x\in F(m)$, whereas $\lE_1$ sends 
any map of $L$-structured cospans as in \cref{eq:2morphismLCspintaF} to the
map of $F$-decorated cospans given by
\[
\begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"k"] & b\ar[l,"o"']\ar[d,"g"] \\
a'\ar[r,"i'"']& m' & b'\ar[l,"o'"]
\end{tikzcd}
\]
along with the morphism of decorations $\tau\maps F(k)(x) \to x'$ in $F(m')$ coming from the middle arrow of \cref{eq:2morphismLCspintaF}. We then make $\lE$ into double functor as in \cref{def:doublefun}. First, both $\lE_0$ and $\lE_1$ are functors that appropriately commute with sources and targets.  Second, we equip $\lE$ with globular 2-isomorphisms $\lE(N) \odot \lE(M) \simrightarrow\lE(N \odot M)$
for any pair of composable horizontal 1-cells $M=\cspn{(a,\bot_a)}{(m,x)}{(b,\bot_b)}{}{}$ and $N=\cspn{(b,\bot_b)}{(n,y)}{(c,\bot_c)}{}{}$ and any object $m$ of $_L\lCsp(\inta F)$. Explicitly, $\mathbb{E}(N)\odot \mathbb{E}(M)$ is given as in \cref{eq:dcospanscomposition} via the pushout cospan and decoration
\begin{equation}\label{eq:1}
 \begin{tikzcd}
  & m+_bn & \\
  a\ar[ur] && b\ar[ul]
 \end{tikzcd}\qquad
  \one\xrightarrow{x\times y}F(m)\times F(n)\xrightarrow{\phi_{m,n}}F(m+n)\xrightarrow{F(\psi)}F(m+_bn).
\end{equation}
On the other hand, $N \odot M$ computed as in \cref{eq:globalpushout} is 
\begin{displaymath}
 \begin{tikzcd}[column sep={.7in,between origins}]
 && (m+_b n,F(v_m)x+_{(\bot_{m+_bn})}F(v_n)y) && \\
 & (m,x) \ar[ur] && (n,y)\ar[ul] \\
 (a,\bot_a)\ar[ur] && (b,\bot_b)\ar[ur]\ar[ul] && (c,\bot_c)\ar[ul]
 \end{tikzcd}
\end{displaymath}
where $v_m,v_n$ are the canonical maps into the pushout.  Under $\lE$ this is mapped to the cospan $\cspn{a}{m+_bn}{b}{}{}$ with decoration $F(v_m)x+F(v_n)y$ in the fiber $F(m+_bn)$ -- since a pushout over an initial object is just a sum.  This decoration is canonically isomorphic to that of \cref{eq:1} due to the following diagram 
\begin{equation}\label{eq:isomorphismdiag}
 \begin{tikzcd}[sep=.4in]
F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[d,"F(v_m)\times F(v_n)"']\ar[dr,phantom,"\stackrel{\phi_{v_m,v_n}}{\cong}"] & F(m+n)\ar[r,"F(\psi)"]\ar[dr,phantom,"\cong"']\ar[d,"F(v_m+v_n)"] & F(m+_bn)\ar[d,"F(1)"]\\
F(m+_bn)\times F(m+_bn)\ar[r,"\phi_{m+_bn,m+_bn}"'] & F\left((m+_bn)+(m+_bn)\right)\ar[r,"F(\nabla)"'] & F(m+_bn)
 \end{tikzcd}
\end{equation}
where the composite of the bottom two arrows is the coproduct $+ \maps F(m+_bn) \times F(m+_bn) \to F(m+_bn)$ as in \cref{eq:explicitstructure1}. The left-hand isomorphism comes from pseudonaturality of $\phi$ as in \cref{eq:pseudonaturality}, whereas the right-hand isomorphism follows from universal properties and pseudofunctoriality of $F$.  Thus, we have a globular 2-isomorphism  $\lE(N) \odot \lE(M) \simrightarrow\lE(N \odot M)$. Notice that since the target decoration is part of the above pushout in the Grothendieck category $\inta F$, this isomorphism we construced is the canonical such that it commutes with the respective cocones. This is useful when checking associativity for example: there is only one such canonical isomorphism $\lE(P)\odot\lE(N)\odot\lE(M)\to\lE(P\odot N\odot M)$ for any three composable structured cospans.

Finally, we construct a globular 2-isomorphism $U_{\lE (m)} \simrightarrow\lE(U_m)$. The identity horizontal 1-cell $U_m$ in $_L\lCsp(\inta F)$ is $\cspn{(m,\bot_m)}{(m,\bot_m)}{(m,\bot_m)}{}{}$ with $1_m$ as the $A$-component and isomorphisms between initial objects in the fibers. Thus $\lE(U_m)$ is the identity cospan in $A$ together with the `initial decoration' $\bot_m\in F(m)$, and so is $U_{\mathbb{E}(m)}$ as in \cref{eq:UFCsp}.

@@

Since both double categories are symmetric monoidal under the running assumptions (\cref{DC,thm:SC}), we now show that this double isomorphism is a symmetric monoidal one, namely that $\lE=(\lE_0,\lE_1)$ is a symmetric monoidal double functor as per \cref{defn:monoidal_double_functor}. Indeed, $\lE_0= \id_\A$ is a symmetric monoidal functor trivially. For $\lE_1$, given two structured cospans $M=(a,\bot_a)\to(m,x)\leftarrow(b,\bot_b)$ and $M'=(a',\bot_{a'})\to(m',x')\leftarrow(b',\bot_{b'})$ in $_L\lCsp(\inta F)$, we compute the appropriate decorated cospans
\begin{align*}
\lE(M)\ot\lE(M')&\;\stackrel{\cref{eq:tensordecoration}}{=}\;a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\phi_{m,m'}(x,x')\in F(m+m')
%\one\xrightarrow{x\times x'}F(m)\times F(m')\xrightarrow{\phi_{m,m'}}F(m+m')
\\
\lE(M\ot M')&\;\stackrel{\cref{eq:tensorstuctured}}{=}\;\lE\Big((a+a',\bot_{a+a'})\to\underbrace{(m+m',F(\iota_m)x+F(\iota_{m'})x')}_{\cref{eq:fibrewisecoprod}}\leftarrow(b+b',\bot_{b+b'})\Big) \\
&=a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }F(\iota_m)x+F(\iota_{m'})x'\in F(m+m')
\end{align*}
The cospan part is equal, whereas we have an isomorphism between the decorations, $F(\iota_m)x+F(\iota_{m'})x'\cong \phi_{m,m'}(x,x')$, given in an analogous way to \cref{eq:isomorphismdiag} as follows:
\begin{displaymath}
\begin{tikzcd}[sep=.4in]
F(m)\times F(m')\ar[r,"\phi_{m,m'}"]\ar[d,"F(\iota_m)\times F(\iota_{m'})"']
\ar[dr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{\iota_m,\iota_n}}{\cong}}"] & F(m+m')\ar[dr,"F(1)"]\ar[d,bend left=60, phantom,"\cong"']\ar[d,"F(\iota_m+\iota_{m'})"']  & \\
F(m+m')\times F(m+m')\ar[r,"\phi_{m+m',m+m'}"'] & F\left((m+m')+(m+m')\right)\ar[r,"F(\nabla)"'] & F(m+m')
 \end{tikzcd}
\end{displaymath}
where the composite of the bottom two arrows is the coproduct operation in $F(m+m')$.
We thus have constructed an isomorphism $\lE(M)\otimes\lE(M') \simrightarrow \lE(M\otimes M')$ for each $M,M'$. 
Again, notice that since the target decoration is part of the sum $\left(m+m',F(\iota_m)x+F(\iota_{m'})x'\right)$ in the Grothendieck category $\inta F$ by \cref{eq:fibrewisecoprod}, this isomorphism we constructured is the canonical such that it commutes with the respective inclusions into the sum, and there is only one such canonical isomorphism $\lE(M)\otimes\lE(M')\otimes\lE(M'')\simrightarrow\lE(M\ot M'\ot M'')$ for any three structured cospans, proving associativity.

@@

\begin{proof}
This follows by applying Shulman's result (\cref{Shulhorizontalbicat}) to the fibrant symmetric monoidal double category $F\lCsp$.
\end{proof}

@@

\begin{thm} \label{thm:bicat_equiv}
Suppose $\A$ has finite colimits and $F \maps(\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor that factors through $\Rex$ as an ordinary pseudofunctor.    Define the symmetric monoidal bicategory $_L\bCsp(\inta F)$ as in \cref{thm:equiv}.   Then there is an isomorphism of symmetric monoidal bicategories
\[      F\bCsp \cong {}_L \bCsp(\inta F)   \]
and of symmetric monoidal categories
\[      F\Csp \cong {}_L \Csp(\inta F)  . \]
\end{thm}

\begin{proof} 
Hansen and Shulman \cite{HS} showed that the passage from symmetric monoidal double categories to symmetric monoidal bicategories is functorial in a suitable sense.  This implies that an isomorphism of symmetric monoidal double categories $\lD \cong \lD'$ gives an isomorphism of symmetric monoidal bicategories $\bD \cong \bD'$.    Since the process of decategorifying a bicategory merely discards 2-morphisms and takes isomorphism classes of 1-morphisms, the isomorphism of symmetric monoidal bicategories $\bD \cong \bD'$ in turn induces an isomorphism of symmetric monoidal categories $\D \cong \D'$.   Thus, the theorem follows from \cref{thm:equiv}. \end{proof}

@@

We thus have two symmetric monoidal double categories: ${}_L \lCsp(\Graph)$ obtained from structured cospans and $F\lCsp$ obtained from decorated cospans. Each of these double categories has $\Fin\Set$ as its category of objects, open graphs as horizontal 1-cells, and maps of open graphs as 2-morphisms.   This suggests that ${}_L \lCsp(\Graph)$  and $F\lCsp$ are isomorphic as symmetric monoidal double categories---and indeed this is true.   This follows from \cref{thm:equiv} together with the fact that $U \maps \inta F \to \Fin\Set$ and $R \maps \Graph \to \Fin\Set$ are isomorphic as symmetric monoidal opfibrations.  %%% Christina: I don't understand why the last part of this sentence is needed. The isomorphism should be obtained straight away from our theorem no? I *kinda* see the point but it looks like a technicality that it could be more obscuring than clarifying...

@@

Given a Petri net with rates $P$ and defining $v$ as above, this system of first-order ordinary differential equations for a function $c \maps \R \to \R^S$:
\[    \frac{d}{dt} c(t) = v(c(t))   \]
is called the \define{rate equation}.    When $P$ is part of an open Petri net with rates
\[
\begin{tikzpicture}[scale=1.5]
\node (D) at (-3,0) {$L(X)$};
\node (E) at (-2,0) {$P$};
\node (F) at (-1,0) {$L(Y)$,};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$i$} (E)
(F) edge node [above] {$o$} (E);
\end{tikzpicture}
\]
we get an open dynamical system called the \define{open rate equation}:
\[     \frac{d}{dt} c(t) = v(c(t))  + i_*(I(t)) - o_*(O(t)) \]
where $I \maps \R \to \R^X$ and $O \maps \R \to \R^X$ are arbitrary smooth functions describing \define{inflows} and \define{outflows}, respectively.   Applying this prescription to the open Petri net with rates shown at the start of this section one gets the differential equations \cref{eq:openPetrir}.  Other examples are worked out in \cite{BP}. %%% Christina: is this where the description of the black box double functor ends? Should we add a finishing line to that end?

We now show that decorated cospan double category $D \lCsp$ of open dynamical systems is not isomorphic to a structured cospan double category via \cref{thm:equiv}.  Recall that in that theorem we start with the data required to build a decorated cospan category, namely a symmetric lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat,\times)$, and show that under certain conditions the resulting opfibration $U \maps \X = \inta F \to \A$ has a left adjoint $L \maps \A \to \X$.   We then obtain an isomorphism between decorated and structured cospan double categories, $F \lCsp \cong {}_L \Csp(\X)$.   We now show that in the case at hand, where $F = D$,  the opfibration $U$ does \emph{not} have a left adjoint.  Thus, the conditions of \cref{thm:equiv} do not hold in this case.

@@

where the unnamed isomorphisms are appropriate composites of the pseudoassociativity of $F$ \cref{eq:omega}. The fact that the two pasted 2-isomorphisms are equal now follows from lengthy pasted diagram calculations using the axioms of a sylleptic lax monoidal pseudofunctor \cite[Definition 16]{DS}.
